Limited plasticity but increased variance in physiological rates across ectotherm populations under climate change

### 2024-07-14 ms.qmd
Adding some general comparisons across major category types, written lnVR results, adding to lnRR and fixed a bug in the get_lifestage function. Need to remember that we're using a different methods to caluclate predictions and this results in posterior being in columns not rows, so flipped this around.

We compiled literature on ectothermic animals that measured
physiological rates (e.g., metabolic rate) at two or more temperatures
after having been acclimated (or acclimatized) at these temperatures. We
used data from a previous meta-analysis [@Seebacher2015-wi] and updated
@Seebacher2015-wi's data by extracting data from suitable studies from
our own searches that followed the same search protocol. For full
details on the search protocol see the *Supplementary Materials*, where
we also provide a PRISMA flow diagram of our extraction process
(@fig-prisma).

@@

### *Climate Data*

Using the thermal time-series data for each location we calculated
metrics of thermal variability across months and years as well as
estimates of thermal predictability (i.e., autocorrelation). To estimate
thermal variability, we calculated the coefficient of variation
($\frac{SD}{M}$, where SD = standard deviation in temperature and M =
the mean temperature for each year). We also estimated thermal
predictability, by calculating the auto-regressive time lag across
months (i.e., a measure of how correlated temperatures were between
months), however, identifying biologically relevant lags for such
diverse taxa is challenging. As such, we present a coarse analysis using
this metric of thermal predictability in the *Supplemental Materials*.

@@

In addition to acclimation, all our models corrected for the possibility
of the log-linearity assumption of our effect sizes being violated
([@fig-concept]; and see Supplementary Materials @fig-lnrrq10_temp). We
predicted that, if $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ were not
strictly log-linear than we would see a decrease in average effect size
for studies applying higher temperature treatments because these
temperatures are expected to either converge on or cross the thermal
optima of the performance curve causing reaction rates to decelerate or
decrease beyond $T_{op}$ [@michaletz2024hotter]. The benefit of such an
approach means that we could still take advantage of the excellent
properties of $Q_{10}$ as an effect size while statistically correcting
for the potential non-linearity that would be expected in the data at
high treatment temperatures. Given that our data included a wide range
of species and habitats, we also included a random slope of maximum
temperature that varied across species because we, *a priori*, expected
that each species will vary in their thermal performance curve and the
chosen temperatures experimentally manipulated along the performance
curve. To improve model interpretation, we also mean-centered the
maximum temperature and included it in our models.

@@

We first tested the degree to which the difference in acute and
acclimation $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ effects varied by
habitat type (i.e., terrestrial, freshwater, and marine). Models
included an interaction between effect type (i.e., acute or acclimation)
and habitat (referred to as “Model 2”). Reduced mean
$lnRR_{Q_{10_{acclimation}}}$ relative to $lnRR_{Q_{10_{acute}}}$
indicates that acclimation to thermal environments results in partial
compensation of physiological rates (i.e., phenotypic plasticity),
whereas no differences between $lnRR_{Q_{10_{acute}}}$ and
$lnRR_{Q_{10_{acclimation}}}$ suggests organisms are not capable of
physiological plasticity [@Havird2020-mj; @Seebacher2015-wi]. In
contrast, a difference in $lnCVR_{Q_{10_{acclimation}}}$ relative to
$lnCVR_{Q_{10_{acute}}}$ would imply that changes in between individual
variation in physiological rates across 10°C differ depending on whether
acute or acclimation responses are measured.

Second, we tested whether acute and acclimation $lnRR_{Q_{10}}$ and
$lnCVR_{Q_{10}}$ differed between whole-organism versus biochemical
traits across habitats by fitting an model with an interaction between
type, habitat and trait category (referred to as "Model 3"). We also
explored how acute and acclimation effects varied across more detailed
trait categories, and these results can be found in the *Supplementary
Materials*. Given that residual variance appeared to differ across trait
categories we fit heteroscedastic residual variance models which were
better supported in all cases ($\Delta AIC_{c}$; marine = 58, freshwater
= 120, and terrestrial = 12). We expected that whole-organism traits
would be more likely to maintain relative variation in physiological
function and be less likely to acclimate because whole-organism function
relies on a greater number of biochemical reactions each with different
thermal sensitivies [@fields2001protein; @angilletta2009thermal].

@@

Finally, we tested whether the change in $lnRR_{Q_{10_{acclimation}}}$
and $lnCVR_{Q_{10}}$ were predicted by a populations climate variability
(CV). We only used $lnRR_{Q_{10_{acclimation}}}$ for these models
because our predictions were specifically focused on the plastic
adjustments to physiological rates. We used all $lnCVR_{Q_{10}}$ data
because there were no differences between $lnCVR_{Q_{10_{acclimation}}}$
and $lnCVR_{Q_{10_{acute}}}$ and we had no *a priori* reason to believe
climate would impact acclimation and acute $lnCVR_{Q_{10}}$ differently.
To test how climate variability impacted mean effect size in wild
populations we fit models that included an interaction between habitat
type and thermal coefficient of variability (CV) as moderators while
controlling for acclimation time and maximum temperature used in the
experiments (referred to as "Model 5"). We also explored how
environmental predictability, as measured by how correlated successive
temperatures at a given location were across months, might also explain
capacity for acclimation. However, such analyses are challenging to
interpret because the temporal scale that is biologically relevant to
different organisms will be different making the choice of lag to
estimate the correlation difficult to apply across taxa. As such, we
report on these results for a single lag (1 month) in the *Supplementary
Materials* but note they do not differ from CV analyses.

@@

To understand the consequences of human-induced climate change on the
potential to impact relative variance in physiological rates we fit a
model that included an interaction between acclimation type, habitat
type, latitude and longitude (referred to as "Model 6"). We assumed that
any change in $lnCVR_{Q_{10}}$ across latitude and longitude could vary
by habitat type (i.e., an interaction between habitat). We used
non-linear tensors for latitude and longitude as any response could be
complicated by local factors (e.g., altitude). Our model included random
effects of species, trait, phylogeny and study. We predicted the
expected change in $lnCVR_{Q_{10}}$ for each wild population in our
dataset at the specific populations latitude and longitude. To do this,
we first converted the predicted $lnCVR_{Q_{10}}$ to a 1°C change as
opposed to 10°C as follows:

@@

@@

### *Performance Curve Simulation*
To better understand the characteristics of the performance curves in a sample that would lead to observed changes in variance across temperature we conducted a simple simulation. To simulate performance curves, we used a asymmetrical Gaussian function (@eq-sim):

@@

where $T$ is the temperature gradient, $\delta$ is the optimal temperature (the temperature where performance is maximized), $\sigma$ the performance breadth, and $\alpha$ the skewness of the performance function or rate variation (see *Supplementary Materials* showing curves).  While we could have used other performance curve functions (e.g., Arrhenius), our intentions were not to model the exact curves (which are not know) but rather identify what might vary across the major habitat types that could contribute to the different patterns of variance change at higher temperatures we observed across habitats.

@@

Capacity for acclimation depended on the habitat. Ectotherms in marine
and freshwater environments showed partial compensation of physiological
rates ([@fig-2]A) amounting to reduced $lnRR_{Q_{10_{acclimation}}}$ of
`r abs(mean(change_f))`% (95% CI:
`r quantile(change_f, c(0.025, 0.975))[1]` to
`r quantile(change_f, c(0.025, 0.975))[2]`) in freshwater and
`r abs(mean(change_m))`% (95% CI:
`r quantile(change_m, c(0.025, 0.975))[1]` to
`r quantile(change_m, c(0.025, 0.975))[2]`) in marine environments. In
contrast, terrestrial ectotherms showed no acclimation (possibly even
inverse acclimation) -- showing a `r abs(mean(change_t))`% increase in
$lnRR_{Q_{10_{acclimation}}}$ (95% CI:
`r quantile(change_t, c(0.025, 0.975))[1]` to
`r quantile(change_t, c(0.025, 0.975))[2]`, [@fig-2]B). All results are
from Model 2 (see “Meta-Analysis” above).

Across habitats the extent to which whole-organism versus biochemical
processes acclimated varied ("Model 3"; [@fig-btrait]A-C; and see
*Supplementary Materials*, @fig-lnrrtrait for more detailed trait analysis). Biochemical
processes were more effective at acclimating compared to whole-organism
traits for ectotherms occupying marine habitats ([@fig-btrait]B),
whereas both whole-organism and biochemical traits acclimated in
freshwater ectotherms ([@fig-btrait]C). Unsurprisingly, neither broad
trait category showed evidence of acclimation in terrestrial ectotherms
([@fig-btrait]A).

@@

  # Now, lets get contrasts between habitats averaged over type
      f_m_c <- f_VR_mean - m_VR_mean   ; pmcmc(f_m_c); mean(f_m_c)
      f_t_c <- t_VR_mean_ef - f_VR_mean; pmcmc(f_t_c); mean(f_t_c)
      m_t_c <- t_VR_mean_ef - m_VR_mean; pmcmc(m_t_c); mean(m_t_c)

@@

        # Contrast statistics
            contrast_lnvr_lh_stats <- data.frame( habitat = rep(c("t", "m", "f"), each = 2),
                                                     stage = rep(c("j", "a"), each = 1),
                                                    diff = apply(data.frame(contrast_lnvr_lh_post), 1, mean),
                                                     l95 = apply(data.frame(contrast_lnvr_lh_post), 1, function(x) quantile(x, 0.025)),
                                                     u95 = apply(data.frame(contrast_lnvr_lh_post), 1, function(x) quantile(x, 0.975)),
                                                       p = apply(data.frame(contrast_lnvr_lh_post), 1, function(x) p_value(pmcmc(x))))
```

```{r, fig-2}
#| label: fig-2
#| fig-cap: "A) Mean acute and acclimation $lnRR_{Q_{10}}$ across ectotherms in marine, freshwater, and terrestrial habitats. B) Mean acute and acclimation $lnVR_{Q_{10}}$ across ectotherms in marine, freshwater and terrestrial habitats. For both plots, k = total number of effect size estimates while the numbers in brackets indicate the number of species. Sample sizes are the same for panel A) and B). In panels A) and B), thick bars are 95% confidence intervals (CI) and thin bars 95% prediction intervals (PI). $\\Beta$ values are the contrasts between acute and acclimation means within each habitat. $\\mu$ values are the overall meta-analytic means averaged across acute and acclimation types within each habitat type. In both cases, their 95% CI's are indicated within square brackets. $p_{MCMC}$ values are the posterior probability of the contrast or overall meta-analytic mean being different from zero. For ease of visualisation, all the raw data plotted for both acute and acclimation type effect sizes are presented as circles."
#| warning: false
#| echo: false

@@

### *Variation in physiological rates increases in aquatic but not terrestrial ectotherms*

Variance in physiological rates ($lnVR_{Q_{10}}$) showed an
increase with increasing temperature across all habitat types with less heterogeneity overall ($I^2_{sv}$ = `r mean(tot_i2_VR)`, 95% CI:
`r quantile(tot_i2_VR, 0.025)` to `r quantile(tot_i2_VR, 0.975)`; "Model 1"; see *Supplementary Materials* @fig-s3). However, physiological rate variance increased significantly more in freshwater (freshwater vs terrestrial: $\beta$ = `r mean(f_t_c)`, 95% CI: `r quantile(f_t_c, 0.025)` to `r quantile(f_t_c, 0.975)`, $p_{MCMC}$ = `r p_value(pmcmc(f_t_c))`) and marine ectotherms (marine vs terrestrial: $\beta$ = `r mean(m_t_c)`, 95% CI: `r quantile(m_t_c, 0.025)` to `r quantile(m_t_c, 0.975)`, $p_{MCMC}$ = `r p_value(pmcmc(m_t_c))`) compared to the limited increase in terrestrial ecotherms (@fig-2). There were no differences between
$lnVR_{Q_{10_{acute}}}$ and $lnVR_{Q_{10_{acclimation}}}$. Differences
in variance between habitats were driven by different scaling relationships between mean and variance across habitats (see *Supplementary
Materials*; @tbl-slopes & @fig-mv). There was a `r mean(t_VR_mean_ef)`%
(95% CI: `r quantile(t_VR_mean_ef, 0.025)[1]` to
`r quantile(t_VR_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(t_VR_mean)`) increase in physiological rate variance
for terrestrial ectotherms and a `r mean(m_VR_mean_ef)`% (95% CI:
`r quantile(m_VR_mean_ef, 0.025)[1]` to
`r quantile(m_VR_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(m_VR_mean)`) increase in variation for marine
ectotherms when temperatures increased by 10$^\circ$C. In contrast,
freshwater ectotherms exhibited a `r mean(f_VR_mean_ef)`%, 95% CI: `r quantile(f_VR_mean_ef, 0.025)[1]`
to `r quantile(f_VR_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(f_VR_mean)` increase in variance at high temperatures
(@fig-2). (results from "Model 2"). The relative differences between habitats were similar when considering $lnCVR_{Q_{10}}$ (See *Supplementary Materials*).

```{r, fig-sim2}
#| label: fig-sim2
#| fig-align: center
#| fig-cap: "Expected $lnVR_{Q_{10}}$ when varying $\\sigma_{_b_{\\delta}}$ = {0.01, 2} and  $\\sigma_{_b_{\\sigma}}$ = {0.01, 2} while fixing $\\sigma_{_b_{\\alpha}}$ = 0.01. In all simulations, $\\delta$ = 35, $\\sigma$ = 9, $\\alpha$ = -15, and *n* = 1000 individuals were simulated for each combination of $\\sigma_{_b_{\\delta}}$, $\\sigma_{_b_{\\sigma}}$. The parameter space that match observed mean $lnVR_{Q_{10}}$ for terrestrial (green), marine (orange) and freshwater (blue) ectotherms. Dashed lines indicate the relative differences between the three habitat types when holding one parameter constant."

#ggsave(lnvr_sim_delta_sigma1, width = 9.5, height = 8, filename = "./output/figures/figsim_new2.pdf")

p_sim2 <- image_read("./output/figures/figsim_new.pdf")
p_sim2
```

Across habitats whole-organism versus biochemical processes showed a
similar change in variance in physiological rates
([@fig-btrait]D-F; see *Supplementary Materials*, @fig-vrtrait) for more detailed trait
analysis), except for biochemical processes in ectotherms occupying
freshwater habitats, which showed small, albeit significant increase in
relative variance when acclimated ([@fig-btrait]F). Reduced relative
variance was particularly prominent for resting metabolic rates and
sprint speed although traits differed in whether they exhibited a
reduction in variation in physiological rates at high temperatures (see
*Supplemental Material*, @fig-vrtrait) ("Model 3"). Each life-history stage
exhibited the same pattern of relative variance change in each of the
habitats ("Model 4"; @fig-cvrlh).

@@

############
# Full fig
############
figbtrait <-(p3 | p1 | p2) / (p3_vr_trait | p1_vr_trait | p2_vr_trait) + plot_annotation(tag_levels = "A") 
ggsave(figbtrait, filename = "./output/figures/fig4.png", width = 28, height = 14, device = "png")

figbtrait <- magick::image_read("./output/figures/fig4.png")
figbtrait
```

@@

fig4 <-(p3 | p1 | p2) / (p3_vr | p1_vr | p2_vr) + plot_annotation(tag_levels = "A") 
ggsave(fig4, filename = "./output/figures/fig5.pdf", width = 28, height = 14, device = "pdf")

fig4 <- magick::image_read("./output/figures/fig5.pdf")
fig4
```

@@

```{r, fig-5}
#| label: fig-5
#| fig-align: center
#| fig-cap: "Predicted mean acclimation (thick black line) $lnRR_{Q_{10_{acclim}}}$ (A) and $lnCVR_{Q_{10_{acclim}}}$ (B) as a function of the Thermal Coefficient of Variation (CV) for wild populations across marine, freshwater and terrestrial habitats. Dashed lines indicate 95% confidence intervals and dotted lines indicate 95% prediction intervals. Model slope ($\\beta$) along with the 95% CI and $p_{MCMC}$ values for the slopes are shown for each habitat."

@@

>>>>

We compiled literature on ectothermic animals that measured
physiological rates (e.g., metabolic rate) at two or more temperatures
after having been acclimated (or acclimatized) at these temperatures. We
used data from a previous meta-analysis [@Seebacher2015-wi] and updated
@Seebacher2015-wi's data by extracting data from suitable studies from
our own searches that followed the same search protocol. We extracted data from an extra
`r length(unique(new_papers$record_num))` papers (with a total of
`r dim(new_papers)[1]` effects; a
`r  (length(unique(new_papers$record_num)) / length(unique(old_papers$record_num)))*100`%
increase in the number of published articles). For full
details on the search protocol see the *Supplementary Materials*, where
we also provide a PRISMA flow diagram of our extraction process
(@fig-prisma).

@@

@@

In addition to acclimation, all our models corrected for the possibility
of the log-linearity assumption of our effect sizes being violated
([@fig-concept]; and see Supplementary Materials @fig-lnrrq10_temp). We
predicted that, if $lnRR_{Q_{10}}$ and $lnVR_{Q_{10}}$ were not
strictly log-linear than we would see a decrease in average effect size
for studies applying higher temperature treatments because these
temperatures are expected to either converge on or cross the thermal
optima of the performance curve causing reaction rates to decelerate or
decrease beyond $T_{op}$ [@michaletz2024hotter]. The benefit of such an
approach means that we could still take advantage of the excellent
properties of $Q_{10}$ as an effect size while statistically correcting
for the potential non-linearity that would be expected in the data at
high treatment temperatures. Given that our data included a wide range
of species and habitats, we also included a random slope of maximum
temperature that varied across species because we, *a priori*, expected
that each species will vary in their thermal performance curve and the
chosen temperatures experimentally manipulated along the performance
curve. To improve model interpretation, we also mean-centered the
maximum temperature and included it in our models.

@@

We first tested the degree to which the difference in acute and
acclimation $lnRR_{Q_{10}}$ and $lnVR_{Q_{10}}$ effects varied by
habitat type (i.e., terrestrial, freshwater, and marine). Models
included an interaction between effect type (i.e., acute or acclimation)
and habitat (referred to as “Model 2”). Reduced mean
$lnRR_{Q_{10_{acclimation}}}$ relative to $lnRR_{Q_{10_{acute}}}$
indicates that acclimation to thermal environments results in partial
compensation of physiological rates (i.e., phenotypic plasticity),
whereas no differences between $lnRR_{Q_{10_{acute}}}$ and
$lnRR_{Q_{10_{acclimation}}}$ suggests organisms are not capable of
physiological plasticity [@Havird2020-mj; @Seebacher2015-wi]. In
contrast, a difference in $lnCVR_{Q_{10_{acclimation}}}$ relative to
$lnVR_{Q_{10_{acute}}}$ would imply that changes in between individual
variation in physiological rates across 10°C differ depending on whether
acute or acclimation responses are measured.

Second, we tested whether acute and acclimation $lnRR_{Q_{10}}$ and
$lnVR_{Q_{10}}$ differed between whole-organism versus biochemical
traits across habitats by fitting an model with an interaction between
type, habitat and trait category (referred to as "Model 3"). We also
explored how acute and acclimation effects varied across more detailed
trait categories, and these results can be found in the *Supplementary
Materials*. Given that residual variance appeared to differ across trait
categories we fit heteroscedastic residual variance models which were
better supported in all cases ($\Delta AIC_{c}$; marine = 58, freshwater
= 120, and terrestrial = 12). We expected that whole-organism traits
would be more likely to maintain relative variation in physiological
function and be less likely to acclimate because whole-organism function
relies on a greater number of biochemical reactions each with different
thermal sensitivies [@fields2001protein; @angilletta2009thermal].

@@

Finally, we tested whether the change in $lnRR_{Q_{10_{acclimation}}}$
and $lnVR_{Q_{10}}$ were predicted by a populations climate variability
(CV). We only used $lnRR_{Q_{10_{acclimation}}}$ for these models
because our predictions were specifically focused on the plastic
adjustments to physiological rates. We used all $lnVR_{Q_{10}}$ data
because there were no differences between $lnVR_{Q_{10_{acclimation}}}$
and $lnVR_{Q_{10_{acute}}}$ and we had no *a priori* reason to believe
climate would impact acclimation and acute $lnVR_{Q_{10}}$ differently.
To test how climate variability impacted mean effect size in wild
populations we fit models that included an interaction between habitat
type and thermal coefficient of variability (CV) as moderators while
controlling for acclimation time and maximum temperature used in the
experiments (referred to as "Model 5"). We also explored how
environmental predictability, as measured by how correlated successive
temperatures at a given location were across months, might also explain
capacity for acclimation. However, such analyses are challenging to
interpret because the temporal scale that is biologically relevant to
different organisms will be different making the choice of lag to
estimate the correlation difficult to apply across taxa. As such, we
report on these results for a single lag (1 month) in the *Supplementary
Materials* but note they do not differ from CV analyses.

@@

To understand the consequences of human-induced climate change on the
potential to impact relative variance in physiological rates we fit a
model that included an interaction between acclimation type, habitat
type, latitude and longitude (referred to as "Model 6"). We assumed that
any change in $lnVR_{Q_{10}}$ across latitude and longitude could vary
by habitat type (i.e., an interaction between habitat). We used
non-linear tensors for latitude and longitude as any response could be
complicated by local factors (e.g., altitude). Our model included random
effects of species, trait, phylogeny and study. We predicted the
expected change in $lnVR_{Q_{10}}$ for each wild population in our
dataset at the specific populations latitude and longitude. To do this,
we first converted the predicted $lnVR_{Q_{10}}$ to a 1°C change as
opposed to 10°C as follows:

@@

  # Contrats adult vs juv across habitats regardless of the acute/acclim
               contrast_lnrr_lh_cat_stats <- rbind(get_lifestage(pred, habitat = "t", newdata),
                                                   get_lifestage(pred, habitat = "m", newdata),
                                                   get_lifestage(pred, habitat = "f", newdata))

@@

### *Performance Curve Simulation*
To better understand the characteristics of the performance curves in a sample that would lead to observed changes in variance across temperature we conducted a simple simulation. To simulate performance curves, we used an asymmetrical Gaussian function:

@@

where $T$ is the temperature gradient, $\delta$ is the optimal temperature (the temperature where performance is maximized), $\sigma$ the performance breadth, and $\alpha$ the skewness of the performance function or rate variation (see @fig-sim in *Supplementary Materials* for example curves).  While we could have modelled performance curves differently than @eq-sim (e.g., Arrhenius), our intentions were not to model the exact curves (which are not know) but rather identify what might vary across the major habitat types that could contribute to the different patterns of variance change at higher temperatures we observed across habitats.

@@

Capacity for acclimation depended on the habitat. Ectotherms in marine
and freshwater environments showed partial compensation of physiological
rates ([@fig-2]A) amounting to reduced $lnRR_{Q_{10_{acclimation}}}$ of
`r abs(mean(change_f))`% (95% CI:
`r quantile(change_f, c(0.025, 0.975))[1]` to
`r quantile(change_f, c(0.025, 0.975))[2]`) in freshwater and
`r abs(mean(change_m))`% (95% CI:
`r quantile(change_m, c(0.025, 0.975))[1]` to
`r quantile(change_m, c(0.025, 0.975))[2]`) in marine environments. In
contrast, terrestrial ectotherms showed no acclimation (possibly even
inverse acclimation) -- showing a `r abs(mean(change_t))`% increase in
$lnRR_{Q_{10_{acclimation}}}$ (95% CI:
`r quantile(change_t, c(0.025, 0.975))[1]` to
`r quantile(change_t, c(0.025, 0.975))[2]`, [@fig-2]A). All results are
from Model 2 (see “Meta-Analysis” above).

Across habitats the extent to which whole-organism versus biochemical
processes acclimated varied ("Model 3"; [@fig-btrait]A-C; and see
*Supplementary Materials*, @fig-lnrrtrait for more detailed trait analysis). Biochemical
processes were more effective at acclimating compared to whole-organism
traits for ectotherms occupying marine habitats ([@fig-btrait]B),
whereas both whole-organism and biochemical traits acclimated in
freshwater ectotherms ([@fig-btrait]C). Unsurprisingly, neither
trait category showed evidence of acclimation in terrestrial ectotherms
([@fig-btrait]A).

@@

  # Now, lets get contrasts between habitats for acute and acclimation. Keep seperate because the challenge is the rates acclimate so adds quite a bif of variability
      f_m_c_a <- f_acute_VR - m_acute_VR; pmcmc(f_m_c_a); mean(f_m_c_a)
      f_t_c_a <- f_acute_VR - t_acute_VR; pmcmc(f_t_c_a); mean(f_t_c_a)
      m_t_c_a <- m_acute_VR - t_acute_VR; pmcmc(m_t_c_a); mean(m_t_c_a)

      f_m_c_aclim <- f_acclim_VR - m_acclim_VR; pmcmc(f_m_c_aclim); mean(f_m_c_aclim)
      f_t_c_aclim <- f_acclim_VR - t_acclim_VR; pmcmc(f_t_c_aclim); mean(f_t_c_aclim)
      m_t_c_aclim <- m_acclim_VR - t_acclim_VR; pmcmc(m_t_c_aclim); mean(m_t_c_aclim)

@@

        # Contrast statistics
            contrast_lnvr_lh_stats <- data.frame( habitat = rep(c("t", "m", "f"), each = 2),
                                                     stage = rep(c("j", "a"), each = 1),
                                                    diff = apply(data.frame(contrast_lnvr_lh_post), 1, mean),
                                                     l95 = apply(data.frame(contrast_lnvr_lh_post), 1, function(x) quantile(x, 0.025)),
                                                     u95 = apply(data.frame(contrast_lnvr_lh_post), 1, function(x) quantile(x, 0.975)),
                                                       p = apply(data.frame(contrast_lnvr_lh_post), 1, function(x) p_value(pmcmc(x))))

        # Contrasts for pooled acute and acclimation across life stage categories. Checked freshwater and looks correct.
            contrast_lnvr_lh_cat_stats <- rbind(get_lifestage(pred_lh_vr, habitat = "t", lh_b_data),
                                                get_lifestage(pred_lh_vr, habitat = "m", lh_b_data),
                                                get_lifestage(pred_lh_vr, habitat = "f", lh_b_data))
```

```{r, fig-2}
#| label: fig-2
#| fig-cap: "(A) Mean acute and acclimation $lnRR_{Q_{10}}$ across ectotherms in marine, freshwater, and terrestrial habitats. (B) Mean acute and acclimation $lnVR_{Q_{10}}$ across ectotherms in marine, freshwater and terrestrial habitats. For both plots, k = total number of effect size estimates while the numbers in brackets indicate the number of species. Sample sizes are the same for panel A and B. In both panels, thick bars are 95% confidence intervals (CI) and thin bars 95% prediction intervals (PI). $\\Beta$ values are the contrasts between acute and acclimation means within each habitat. $\\mu$ values are the overall meta-analytic means averaged across acute and acclimation types within each habitat type. In both cases, their 95% CI's are indicated within square brackets. $p_{MCMC}$ values are the posterior probability of the contrast or overall meta-analytic mean being different from zero. For ease of visualisation, all the raw data plotted for both acute and acclimation type effect sizes are presented as circles."
#| warning: false
#| echo: false

@@

@@

############
# Full fig
############
figbtrait <-(p3 | p1 | p2) / (p3_vr_trait | p1_vr_trait | p2_vr_trait) + plot_annotation(tag_levels = "A") 
ggsave(figbtrait, filename = "./output/figures/fig5.png", width = 28, height = 14, device = "png")

figbtrait <- magick::image_read("./output/figures/fig5.png")
figbtrait
```

@@

fig4 <-(p3 | p1 | p2) / (p3_vr | p1_vr | p2_vr) + plot_annotation(tag_levels = "A") 
ggsave(fig4, filename = "./output/figures/fig6.pdf", width = 28, height = 14, device = "pdf")

fig4 <- magick::image_read("./output/figures/fig6.pdf")
fig4
```

### *Variation in physiological rates increases but to a greater extent in aquatic compared terrestrial ectotherms*

Variance in physiological rates ($lnVR_{Q_{10}}$) showed an
increase with increasing temperature across all habitat types ([@fig-2]B) with less heterogeneity overall ($I^2_{sv}$ = `r mean(tot_i2_VR)`, 95% CI:
`r quantile(tot_i2_VR, 0.025)` to `r quantile(tot_i2_VR, 0.975)`; "Model 1"; see *Supplementary Materials* @fig-s3). However, physiological rate variance increased significantly more in freshwater for acute resposnes (freshwater vs terrestrial: $\beta$ = `r mean(f_m_c_a)`, 95% CI: `r quantile(f_m_c_a, 0.025)` to `r quantile(f_m_c_a, 0.975)`, $p_{MCMC}$ = `r p_value(pmcmc(f_m_c_a))`), but not for acclimation responses because increases in rates were dampended by acclimation resulting in smaller increases in variance (freshwater vs terrestrial: $\beta$ = `r mean(f_m_c_aclim)`, 95% CI: `r quantile(f_m_c_aclim, 0.025)` to `r quantile(f_m_c_aclim, 0.975)`, $p_{MCMC}$ = `r p_value(pmcmc(f_m_c_aclim))`). While marine ectotherms had larger increases in physiological rate variance compared to terrestrial ectoterms these were not signigicant (Acute: $\beta$ = `r mean(m_t_c_a)`, 95% CI: `r quantile(m_t_c_a, 0.025)` to `r quantile(m_t_c_a, 0.975)`, $p_{MCMC}$ = `r p_value(pmcmc(m_t_c_a))`; Acclimation: $\beta$ = `r mean(m_t_c_aclim)`, 95% CI: `r quantile(m_t_c_aclim, 0.025)` to `r quantile(m_t_c_aclim, 0.975)`, $p_{MCMC}$ = `r p_value(pmcmc(m_t_c_aclim))`)([@fig-2]B). Marine and freshwater did not differ in how much variance changed at higher temperatures (Acute: $\beta$ = `r mean(f_m_c_a)`, 95% CI: `r quantile(f_m_c_a, 0.025)` to `r quantile(f_m_c_a, 0.975)`, $p_{MCMC}$ = `r p_value(pmcmc(f_m_c_a))`; Acclimation: $\beta$ = `r mean(f_m_c_aclim)`, 95% CI: `r quantile(f_m_c_aclim, 0.025)` to `r quantile(f_m_c_aclim, 0.975)`, $p_{MCMC}$ = `r p_value(pmcmc(f_m_c_aclim))`). There were no differences between
$lnVR_{Q_{10_{acute}}}$ and $lnVR_{Q_{10_{acclimation}}}$ in any habitat ([@fig-2]B). Differences
in variance between habitats were driven by different scaling relationships between mean and variance with slopes varying across habitat types (see *Supplementary
Materials* analyses with $lnCVR_{Q_{10}}$; @tbl-slopes & @fig-mv).

Overall, there was a `r mean(t_VR_mean_ef)`%
(95% CI: `r quantile(t_VR_mean_ef, 0.025)[1]` to
`r quantile(t_VR_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(t_VR_mean)`) increase in physiological rate variance
for terrestrial ectotherms and a `r mean(m_VR_mean_ef)`% (95% CI:
`r quantile(m_VR_mean_ef, 0.025)[1]` to
`r quantile(m_VR_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(m_VR_mean)`) increase in variation for marine
ectotherms when temperatures increased by 10$^\circ$C. In contrast,
freshwater ectotherms exhibited a `r mean(f_VR_mean_ef)`%, 95% CI: `r quantile(f_VR_mean_ef, 0.025)[1]`
to `r quantile(f_VR_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(f_VR_mean)` increase in variance at high temperatures
(@fig-2). (results from "Model 2"). The relative differences between habitats were similar when considering $lnCVR_{Q_{10}}$ (See *Supplementary Materials*).

```{r, fig-sim2}
#| label: fig-sim2
#| fig-align: center
#| fig-cap: "Expected $lnVR_{Q_{10}}$ when varying individual variation in thermal breadth($\\sigma_{_b_{\\delta}}$ = {0.01, 2}) and  thermal optima ($\\sigma_{_b_{\\sigma}}$ = {0.01, 2}) while fixing the rate variance constant ($\\sigma_{_b_{\\alpha}}$ = 0.01). In all simulations, $\\delta$ = 35, $\\sigma$ = 9, $\\alpha$ = -15, and *n* = 1000 individuals were simulated for each combination of $\\sigma_{_b_{\\delta}}$ and $\\sigma_{_b_{\\sigma}}$. The parameter space that matches the observed mean $lnVR_{Q_{10}}$ for terrestrial (green), marine (orange) and freshwater (blue) ectotherms is also shown. Dashed lines indicate the relative differences between the three habitat types when holding one variance parameter constant."

#ggsave(lnvr_sim_delta_sigma1, width = 9.5, height = 8, filename = "./output/figures/fig4.pdf")

p_sim2 <- image_read("./output/figures/fig4_new.pdf")
p_sim2
```

Across habitats biochemical processes tended to result in greater increases in variance at higher temperatures, but this was only significant for marine ectotherms (biochemical/whole-organism contrasts: Marine: $\beta$ = `r contrast_lnvr_trait_cat_stats[2,3]`, 95% CI: `r contrast_lnvr_trait_cat_stats[2,4]` to `r contrast_lnvr_trait_cat_stats[2,5]`, $p_{MCMC}$ = `r contrast_lnvr_trait_cat_stats[2,6]`; Freshwater: $\beta$ = `r contrast_lnvr_trait_cat_stats[3,3]`, 95% CI: `r contrast_lnvr_trait_cat_stats[3,4]` to `r contrast_lnvr_trait_cat_stats[3,5]`, $p_{MCMC}$ = `r contrast_lnvr_trait_cat_stats[3,6]`; Terrestrial: $\beta$ = `r contrast_lnvr_trait_cat_stats[1,3]`, 95% CI: `r contrast_lnvr_trait_cat_stats[1,4]` to `r contrast_lnvr_trait_cat_stats[1,5]`, $p_{MCMC}$ = `r contrast_lnvr_trait_cat_stats[1,6]`)
([@fig-btrait]D-F; see also  *Supplementary Materials*, @fig-vrtrait for more detailed trait
analysis)("Model 3"). Interestingly, variance increases for biochemical traits is reduced during acclimation in marine ectotherms ([@fig-btrait]E).

Each life-history stage exhibited the same pattern of variance change in each of the
habitats (Adult-Juvenile contrasts: Marine: $\beta$ = `r contrast_lnvr_lh_cat_stats[2,3]`, 95% CI: `r contrast_lnvr_lh_cat_stats[2,4]` to `r contrast_lnvr_lh_cat_stats[2,5]`, $p_{MCMC}$ = `r contrast_lnvr_lh_cat_stats[2,6]`; Freshwater: $\beta$ = `r contrast_lnvr_lh_cat_stats[3,3]`, 95% CI: `r contrast_lnvr_lh_cat_stats[3,4]` to `r contrast_lnvr_lh_cat_stats[3,5]`, $p_{MCMC}$ = `r contrast_lnvr_lh_cat_stats[3,6]`; Terrestrial: $\beta$ = `r contrast_lnvr_lh_cat_stats[1,3]`, 95% CI: `r contrast_lnvr_lh_cat_stats[1,4]` to `r contrast_lnvr_lh_cat_stats[1,5]`, $p_{MCMC}$ = `r contrast_lnvr_lh_cat_stats[1,6]`), with no differences between acute and acclimation effect types ("Model 4"; @fig-cvrlh).

@@

```{r, fig-5}
#| label: fig-5
#| fig-align: center
#| fig-cap: "Predicted mean acclimation (thick black line) $lnRR_{Q_{10_{acclim}}}$ (A-C) and $lnVR_{Q_{10_{acclim}}}$ (E-G) as a function of the Thermal Coefficient of Variation (CV) for wild populations across marine, freshwater and terrestrial habitats. Dashed lines indicate 95% confidence intervals and dotted lines indicate 95% prediction intervals. Model slope ($\\beta$) along with the 95% CI and $p_{MCMC}$ values for the slopes are shown for each habitat."

@@

### *Climate data*
